Edge sign prediction in general graphs and Correlation Clustering

### 2017-10-31 cc/bias_directed.tex
add missing experiments at the beginning of CC chapter

A trivial way to turn an undirected graph $G$ into a directed graph $G'$ is to let $V'=V$ and, for
every edge $(u,v)$ in $E$, to add both \euv{} and \evu{} to $E'$ with the same sign as $(u,v)$. In
terms of our generative model, this corresponds to putting all the probability mass of $\mu$ on the
$p=q$ diagonal. This is clearly not a very satisfactory solution, for it removes one degree of
freedom from the model. To illustrate this point, we conduct the following experiment. We use our
previous \dssn{} datasets and remove the edge direction. As mentioned earlier, for few pairs of
nodes there are reciprocal edges of different signs, in which case we pick a sign arbitrarily. Given
those undirected graphs, we orient them using the approach described above and compare our method
with the \complowrank{} approach ran directly on the undirected graphs. This is a fair comparison,
for \complowrank{} is designed to works solely on undirected graphs.\Todo{run the undirected
experiments.}

>>>>

A trivial way to turn an undirected graph $G$ into a directed graph $G'$ is to let $V'=V$ and, for
every edge $(u,v)$ in $E$, to add both \euv{} and \evu{} to $E'$ with the same sign as $(u,v)$. In
terms of our generative model, this corresponds to putting all the probability mass of $\mu$ on the
$p=q$ diagonal. This is clearly not a very satisfactory solution, for it removes one degree of
freedom from the model. To illustrate this point, we conduct the following experiment. We use our
previous \dssn{} datasets and remove the edge direction. As mentioned earlier, for few pairs of
nodes there are reciprocal edges of different signs, in which case we pick a sign arbitrarily. Given
those undirected graphs, we orient them using the approach described above and compare our method
with the \complowrank{} approach ran directly on the undirected graphs. This is a fair comparison,
for \complowrank{} is designed to works solely on undirected graphs.

As we can see in \autoref{tab:bias_exp_undir}, our methods perform worse than when running on the
original directed graphs. Looking at \uslpropGsec{} on a 15\% training set as an example, we observe
that the MCC decreases by one (\kiw{}) to almost ten (\epi{}) points. On the other hand, the
\complowrank{} seems to perform better. Recall however that because it is an undirected method in
the first place and because we roughly double the number of signed edges, a 15\% training size means
the algorithm gets to observe more signs and therefore perform better. This small experiment thereby
shows that although the performance remains decent, not having direction information hurt our bias.

\input{bias_exp_undir}

### 2017-10-31 cc/bias_model.tex
add some ref
add missing experiments at the beginning of CC chapter

Namely, we consider \emph{gene regulatory networks}. The nodes of such directed graphs are various
chemicals (such as genes, proteins or messenger RNAs) and a directed edge \euv{} indicates that one
node $u$ influences the concentration of another node $v$ through a chemical reaction. This
influence can be positive (that is, an increase of $u$ concentration results in an increase of $v$
concentration) or negative (that is, an increase of $u$ concentration results in a decrease of $v$
concentration). In this context, we could now interpret the \enquote{niceness} of $u$ as it ability
to accelerate chemical reactions, and its \enquote{popularity} as its propensity to take part in
reactions accelerated by other nodes. Intuitively, this is quite far-fetched. To demonstrate this,
we borrow several gene regulatory networks from \autocite[Table 1]{BioSigned09} and perform the same
experiments as in the previous chapter.\Todo{get the data and run the experiments.}

>>>>

Namely, we consider \emph{gene regulatory networks}. The nodes of such directed graphs are various
chemicals (such as genes, proteins or messenger RNAs) and a directed edge \euv{} indicates that one
node $u$ influences the concentration of another node $v$ through a chemical reaction. This
influence can be positive (that is, an increase of $u$ concentration results in an increase of $v$
concentration) or negative (that is, an increase of $u$ concentration results in a decrease of $v$
concentration). In this context, we could now interpret the \enquote{niceness} of $u$ as it ability
to accelerate chemical reactions, and its \enquote{popularity} as its propensity to take part in
reactions accelerated by other nodes. Intuitively, this is quite far-fetched. To demonstrate this,
we borrow the following three gene regulatory networks from \autocite[Table 1]{BioSigned09}, and
display their statistics in \autoref{tab:bias_bio_dataset}:
\begin{enumerate}[1.]
  \item The signaling pathways of the \hip{} CA1 neuron~\autocite{Hippocamp05}.
  \item The interactions between genes and their products that are involved in \can{} development in
    humans and mouses~\autocite{Cancer07}.
  \item The \reg{} database\autocite{RegulonDB16}, summarizing the connection between transcription
    factors and their targets in the Escherichia coli K-12 bacteria. We download four
    files\footnote{\url{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}} containing
    experimental evidence and keep only the positive and negative edges with \textsf{Strong}
    support.
\end{enumerate}
%\url{http://causalbionet.com/}

\begin{table}[bt]
  \centering
  \small
  \caption{Biological dataset properties. The columns have the same meaning as in
    \autoref{tab:dataset} \vpageref{tab:dataset}\label{tab:bias_bio_dataset}}
  \begin{tabular}{lrrrrrrrr}
    \toprule
    Dataset & \thead{$|V|$}       & \thead{$|E|$}       & \thead{$\frac{|E|}{|V|}$} &
    \thead{$\frac{|E^+|}{|E|}$} & \thead{$\frac{\Psi^2_{G''}(Y)}{|E|}$} &
    \thead{$\frac{\Psi_G(Y)}{|E|}$} & \thead{reciprocal\\ edges} &
    \thead{reciprocal\\ disagreement} \\
    \midrule                                                                                                   
    \hip{} &  \np{501} & \np{1046} &               2.1 &               69.5\% &                       .056 &                    .124 &       0.2\% &              0.0\% \\
    \can{} & \np{1240} & \np{3065} &               2.5 &               78.4\% &                       .047 &                    .108 &       5.5\% &             27.1\% \\
    \reg{} & \np{1700} & \np{2570} &               1.5 &               50.3\% &                       .060 &                    .148 &       0.0\% &              0.0\% \\
    \bottomrule
  \end{tabular}
\end{table}

\input{bias_exp_bio.tex}

Compared with the \dssn{} of the previous chapter, those biological are smaller, have lower average
degree and little to no reciprocal edges. Trying to solve the \esp{} problem using the same
procedure as in \autoref{s:exp}, we read in \autoref{tab:cc_bio_exp} that our methods have lower
absolute MCC performance (for instance, \uslpropGsec{} with a 15\% training size achieves $20$, $24$
and $45$ compared with values between $36$ and $61$ on \dssn{}).  To further illustrate the mismatch
between the troll bias and biological networks, notice that the standard deviation figures are
several times larger than those reported in \autoref{tab:all_mcc}.
